---
title: "FML A4"
author: "Tanmaiyee"
date: "2024-03-17"
output: html_document
---


# Summary
Exploring and understanding the financial data gathered from 21 firms from the Pharmaceuticals data provided.

## Objective: Using cluster analysis, the required data is explored amd analyzed.


#### Loading the data set.
```{r}
library(readr)
Pharmaceuticals <- read_csv("Pharmaceuticals.csv")
View(Pharmaceuticals)
# The data is starting with two extra columns i.e., "Symbol" & "Name" hence removing them from data set.
Pharma_data <- Pharmaceuticals[,-(1:2)]
```


##### A. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.

```{r}
library(tidyverse) # data manipulation
library(factoextra) # clustering alogirthms & visualization
library(ggplot2)
library(flexclust)
library(cluster)
library(gridExtra)
set.seed(123)
head(Pharma_data)

# Selecting numerical variables 1 to 9
numerical_Var <- Pharma_data[,1:9]

# Scaling the data frame
df <- scale(numerical_Var)


# Performing k-mean with different k values.
k1 <- kmeans(df, centers = 3, nstart = 25)
k1_plot <- fviz_cluster(k1, data = df)+ggtitle("k=3")
k2 <- kmeans(df, centers = 4, nstart = 15)
k2_plot <- fviz_cluster(k2, data = df)+ggtitle("k=4")
k3 <- kmeans(df, centers = 5, nstart = 70)
k3_plot <- fviz_cluster(k3, data = df)+ggtitle("k=5")

grid.arrange(k1_plot,k2_plot,k3_plot, nrow = 3)


distance <- dist(df, method = "euclidean")
fviz_dist(distance)

# finding out the best K value using "Silhouette Method"
fviz_nbclust(df,FUNcluster = kmeans,method = "silhouette")+labs(subtitle="Silhouette Method")

#Running the final K value 5.
set.seed(200)
Best_cluster<- kmeans(df, 5, nstart = 70)
print(Best_cluster)
clusplot(df,Best_cluster$cluster, color = TRUE, labels = 2,lines = 0)
```


##### From the above analysis the best fit for **K value is 5** which is obtained usimg **"Silhouette Method"**.


##### B.Interpret the clusters with respect to the numerical variables used in forming the clusters.

```{r}
PC_Cluster <- Pharma_data[,(10:12)]%>% mutate(clusters = Best_cluster$cluster)%>% arrange(clusters, .desc = TRUE)
```


##### From the above analysis and plots, the rows

  + 1,3,4,7,10,16,19,21 are categorized into Cluster 1;
  + 6, 8, 12 into cluster 2; 
  + 5, 9, 14, 20 into cluster 3; 
  + 2, 18 into cluster 4; 
  + 11, 13, 15, 17 into cluster 5.


##### C. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)

```{r}
library(dplyr)
plot_Med<-ggplot(PC_Cluster, mapping = aes(factor(clusters), fill=Median_Recommendation))+geom_bar(position = 'dodge')+labs(x ='No of clusters')

plot_Loc<- ggplot(PC_Cluster, mapping = aes(factor(clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='No of clusters')

plot_Ex<- ggplot(PC_Cluster, mapping = aes(factor(clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='No of clusters')

grid.arrange(plot_Med, plot_Loc, plot_Ex)  
```


  + **Cluster 1**
It has equal hold and moderate buy for the locations UK and US with NYSE as exchange.
  + **Cluster 2**
It has hold, moderate buy, moderate sell and strong buy of which "Hold has more weightage in the locations UK, US and Switzerland with NYSE as exchange.
  + **Cluster 3**
It has moderate buy and sell in the locations in France,Ireland and US with NYSE as exchange.
  + **Cluster 4**
It has more hold followed by moderate buy for the locations Germany and US with AMEX, NASDAQ and NYSE as exchange.
  + **Cluster 5**
It has equal hold and moderate for Canada and US with NYSE as exchange.


##### D. Provide an appropriate name for each cluster using any or all of the variables in the dataset.

Cluster 1  - Equal Hold-Buy
Cluster 2  - High Hold
Cluster 3  - Buy-Sell
Cluster 4  - More Hold-Buy
Cluster 5  - Low Equal Hold-Buy



